---
author: "Mark Bruner"
date: "12/2/20"
output:
  pdf_document: default
title: "Final Exam"
fig_caption: yes
keep_tex: yes
latex_engine: pdflatex
highlight: haddock
citation_package: natbib
pandoc_args: --listings
template: 'preamble.tex'
---

## **Defining the Problem, Providing the Context, and Discussing Factors**

1. **Problem Definition**
  + **Problem to Solve:** Forming equitable groups for a class project.
  + **Primary Objective:** **Maximize** the chance that a group performs well on a class project.

2. **Class Project Context**
  + **Class:** Business Analytics Course
  + **Class Project:** The groups are asked to create a model to predict whether a customer will churn or not based on a dataset provided by a cellular company. 

3. **Problem Type:** Assignment Problem

**Factors** 

1. **Factors of a Group**
+ **1. Visualization Knowledge** (variable column header: visual_know)
  + The group member will create the visualizations for the report and perform the EDA.

+ **2. Programming Skill** (variable column header: prog_skill)
  + This group member will create the model for predicting churn.

+ **3. Presentation Ability** (variable column header: pres_ability) 
  + This group member will create the presentation as well as give the presentation.

  
2. **Data Collection and Definitions**
  + The professor created a survey to assess a student's comfort level with public speaking. The levels they could choose from are as follows: 
    + 1: Significant fear and not confident
    + 2: Some fear and has adequate confidence
    + 3: Comfortable and confident
    + 4: Enthusiastic, excited, and higly confident
  
  
  
  + The professor will assign a value to each student based upon submitted work on how capable a student is in programming and visualization knowledge since there are an adequate amount of previous assignment that students have completed to base this rating. The values are defined as follow:
    + 1: Poor
    + 2: Adequate
    + 3: Good
    + 4: Excellent
 
3. **Decision Variables**
+ The decision variables represents the possibility of each of the 12 students occupying one of the 3 roles in the 4 groups.
$$
x_{11},~x_{12},~x_{13},...,~x_{ij},~~~~~~~~~~~where~i ={1, 2, 3,...12}~and~j = 1, 2,3,...,12
$$

4. **Objective Function**
+ The objective function represents the ability of each of the 12 students for the 3 roles in the 4 groups. It needs to be maximized since we want the best 3 students combination of their abilities for each group.

$$
maximize:
\sum_{i=1}^{12}\sum_{j=1}^{12} c_{ij}x_{ij}
$$

5. **Constraints**
+ Only one student can occupy one role in one group.
+ Only one role can be assigned to one student in each group.
+ I want to have at least one 4 in each group and at the very least two 3's. I made the sum of each group's total score ">=" 10 to achieve that goal. 
+ All decision variables will be non-negative.

such that
$$
\sum_{j=1}^{12} x_{ij} =1, 
~~~~~~~~~~~~~for~all~i = 1, 2, ..., 12 
$$

$$
\sum_{i=1}^{12} x_{ij} = 1, 
~~~~~~~~~~~~~for~all~j = 1, 2, ..., 12
$$
$$
Groups~Constraints:
$$
$$
Group~1:\sum_{i=1}^{12} c_{ij}x_{ij}\geq10~~~~~~~~~~~~~where~j = 1, 2,~and~3
$$

$$
Group~2:\sum_{i=1}^{12} c_{ij}x_{ij}\geq10~~~~~~~~~~~~~where~j = 4, 5,~and~6
$$

$$
Group~3:\sum_{i=1}^{12} c_{ij}x_{ij}\geq10~~~~~~~~~~~~~where~j = 7, 8,~and~9
$$

$$
Group~4:\sum_{i=1}^{12} c_{ij}x_{ij}\geq10~~~~~~~~~~~~~where~j = 10, 11,~and~12
$$


$$
and ~x_{ij}\geq0. 
$$

```{r}
rm(list=ls())
library(lpSolveAPI)
```
  
  
```{r,results=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=90)}

# Assuring that the data set does not change every time I run the entire code.
set.seed(15)

# Randomly generating the data set using the guidelines from the "Data Collection and Definition" section.
pres_ability <- sample.int(4, 12, replace = TRUE) # Allowing for duplicate numbers 
#in the dataset since multiple students could have the same variable skill, knowledge, 
# or ability.
prog_skill <- sample.int(4, 12, replace = TRUE)
vis_know <- sample.int(4, 12, replace = TRUE) 

# Creating objective function vector for assignment formulation.
group_obj <- c(pres_ability, prog_skill, vis_know, 
               pres_ability, prog_skill, vis_know, 
               pres_ability, prog_skill, vis_know, 
               pres_ability, prog_skill, vis_know)

# creating a df for the group objective function coefficients to be able to use later
# to check the results.
group1 <- as.data.frame(cbind(pres_ability, prog_skill, vis_know))
group2 <- group1
group3 <- group1
group4 <- group1

# Naming the columns and rows.
group <- as.data.frame(as.matrix(cbind(group1, group2, group3, group4), 12,12))
colnames(group) <- c("VisGrp1", "ProgGrp1", "PresGrp1", 
                     "VisGrp2", "ProgGrp2", "PresGrp2", 
                     "VisGrp3", "ProgGrp3", "PresGrp3", 
                     "VisGrp4", "ProgGrp4", "PresGrp4")

rownames(group) <- c("Student1", "Student2", "Student3", 
                     "Student4", "Student5", "Student6", 
                     "Student7", "Student8", "Student9", 
                     "Student10", "Student11", "Student12")
group
```

### Objective Function
```{r,results=FALSE}
# Creating the lprec for this problem.
lprec <- make.lp(0,144)
lp.control(lprec, sense = "max")
set.type(lprec, 1:144, type = c("integer"))
set.objfn(lprec, group_obj)
```

### Constraints
```{r, message = FALSE}
# Only one role per student.
add.constraint(lprec, c(rep(1, 12)), indices = c(1:12), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(13:24), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(25:36), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(37:48), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(49:60), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(61:72), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(73:84), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(85:96), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(97:108), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(109:120), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(121:132), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = c(133:144), "=", 1)

# Only one student per role.
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(1, 13, 25, 37, 49, 61, 73, 85, 97, 109, 121, 133), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(2, 14, 26, 38, 50, 62, 74, 86, 98, 110, 122, 134), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(3, 15, 27, 39, 51, 63, 75, 87, 99, 111, 123, 135), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(4, 16, 28, 40, 52, 64, 76, 88, 100, 112, 124, 136), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(5, 17, 29, 41, 53, 65, 77, 89, 101, 113, 125, 137), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(6, 18, 30, 42, 54, 66, 78, 90, 102, 114, 126, 138), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(7, 19, 31, 43, 55, 67, 79, 91, 103, 115, 127, 139), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(8, 20, 32, 44, 56, 68, 80, 92, 104, 116, 128, 140), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(9, 21, 33, 45, 57, 69, 81, 93, 105, 117, 129, 141), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(10, 22, 34, 46, 58, 70, 82, 94, 106, 118, 130, 142), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(11, 23, 35, 47, 59, 71, 83, 95, 107, 119, 131, 143), "=", 1)
add.constraint(lprec, c(rep(1, 12)), indices = 
                 c(12, 24, 36, 48, 60, 72, 84, 96, 108, 120, 132, 144), "=", 1)

# At least one 4 rating and two 3 ratings per group for each role.
add.constraint(lprec, c(pres_ability, prog_skill, vis_know), 
               indices = 
                 c(1, 13, 25, 37, 49, 61, 73, 85, 97, 109, 121, 133, 
                   2, 14, 26, 38, 50, 62, 74, 86, 98, 110, 122, 134, 
                   3, 15, 27, 39, 51, 63, 75, 87, 99, 111, 123, 135), ">=", 10)
add.constraint(lprec, c(pres_ability, prog_skill, vis_know), 
               indices = c(4, 16, 28, 40, 52, 64, 76, 88, 100, 112, 124, 136, 
                           5, 17, 29, 41, 53, 65, 77, 89, 101, 113, 125, 137, 
                           6, 18, 30, 42, 54, 66, 78, 90, 102, 114, 126, 138), ">=", 10)
add.constraint(lprec, c(pres_ability, prog_skill, vis_know), 
               indices = 
                 c(7, 19, 31, 43, 55, 67, 79, 91, 103, 115, 127, 139, 
                   8, 20, 32, 44, 56, 68, 80, 92, 104, 116, 128, 140, 
                   9, 21, 33, 45, 57, 69, 81, 93, 105, 117, 129, 141), ">=", 10)
add.constraint(lprec, c(pres_ability, prog_skill, vis_know), 
               indices = c(10, 22, 34, 46, 58, 70, 82, 94, 106, 118, 130, 142, 
                           11, 23, 35, 47, 59, 71, 83, 95, 107, 119, 131, 143, 
                           12, 24, 36, 48, 60, 72, 84, 96, 108, 120, 132, 144), ">=", 10)
```


```{r}
write.lp(lprec, filename = "final.lp", type = "lp")
lprec
```

```{r}
# Solve LP model
solve(lprec)
```

```{r,results=FALSE}
# The maximum value for the objective function.
get.objective(lprec)

# Decision Variables
get.variables(lprec) -> assign
solution <- matrix(assign, nrow = 12, ncol = 12, byrow = TRUE)

# Multiplying the original group ratings matrix by the solution group assignments.
group*solution -> group_assign 
group_assign
```

```{r}
# Group member combinations.
group1 <- cbind(group_assign[c(1, 7, 11), 1:3], group[c(1, 7, 11), 1:3])
group1
group2 <- cbind(group_assign[c(2, 8, 6), 4:6], group[c(2, 8, 6), 4:6])
group2
group3 <- cbind(group_assign[c(9,5,3), 7:9], group[c(9,5,3), 7:9])
group3
group4 <- cbind(group_assign[c(12,4,10), 10:12], group[c(12,4,10), 10:12])
group4
```


**Group 1:**
Student 11: Visual Role with skill of 4
Student 1: Programming skill of 4
Student 7: Presentation skill of 2

**Group 2:**
Student 2: Visual Role with skill of 3
Student 8: Programming skill of 3
Student 6: Presentation skill of 4

**Group 3:**
Student 12: Visual Role with skill of 3
Student 5: Programming skill of 3
Student 3: Presentation skill of 4

**Group 4:**
Student 9: Visual Role with skill of 3
Student 4: Programming skill of 4
Student 10: Presentation skill of 3

The above groups represents and equitable split of the class for the project.
Group 1 has two members with 4 but to balance that they have a group member with a 2. 
The rest of the groups have one 4 rating and two 3's which I believe makes the groupings equitable and
achieve the result that I was hoping to accomplish.



